A Variant of G\"{o}del's Ontological Proof \\ in a Natural Deduction Calculus \\ (Draft)

### 2014-09-19 Papers/2014/Universalis/ModalCollapse.tex
Reorganize folders

A crucial step of most ontological arguments is the claim that if
God's existence is possible, then it is necessary.  This is Lemma
\textbf{L2} in G\"odel's proof.  In the modal square of opposition
(Fig. \ref{fig:square}), this is an unusual situation in which the
\textbf{I} corner must imply and entail the \textbf{A} corner, 
in the particular case when $\phi$ is $\exq x G(x)$.  
G\"odel's proof shows that his axioms are indeed 
strong enough to invert the direction of
entailment for this choice of $\phi$.  
This observation, however, immediately leads 
to the question whether the axioms are eventually 
even strong enough to enable the inverted entailment
for any arbitrary sentence $\phi$.  
That is essentially the question asked by 
Sobel \cite{Sobel1987}, and his proof of the modal collapse
(\textbf{MC}, cf. Fig.~\ref{fig:collapse}) 
provides an affirmative answer. It is possible to show
that this form of the modal collapse entails (in modal logic
\textbf{\emph{K}}) a collapse of the modal square (\textbf{MCs}),
causing the subcontraries to entail 
(and even imply) their respective
contraries. Normally, as shown in Fig. \ref{fig:square}, 
in the modal square of opposition only the 
other direction of entailment holds: the
contraries entail their subcontraries, 
assuming the \emph{modal existential import} \textbf{ExImp}
\cite{WhatToCiteHere}.
\footnote{C.: We need a reference here; I am still
  searching for good papers; this here is also interesting:
  \url{http://www.tandfonline.com/doi/pdf/10.1080/01445340.2013.764962}}

@@

Independently of the variant of the ontological argument 
under consideration, the following can be said about 
classes of collapsing propositions:
\begin{enumerate}
\item Valid propositions are collapsing: 
if $\phi$ is valid, then $\collapse{\phi}$ is valid.
%
\item The class of collapsing propositions 
is closed under logical equivalence: 
if $\collapse{\phi}$ is valid and $\phi \biimp \phi'$ is valid, 
then $\collapse{\phi'}$ is valid.
%
\item The class of collapsing propositions is 
not generally closed under equi-validity: 
even if $\collapse{\phi}$ is valid and $\phi$ and $\phi'$ 
are equi-valid, $\collapse{\phi'}$ may not be valid.
%
\item The class of collapsing propositions is 
not generally closed under implication: 
even if $\collapse{\phi}$ is valid and 
$\phi \imp \phi'$ is valid, 
$\collapse{\phi'}$ may not be valid.
\end{enumerate}

>>>>

A crucial step of most ontological arguments is the claim that if
God's existence is possible, then it is necessary.  This is Lemma
\textbf{L2} in G\"odel's proof.  In the modal square of opposition
(Fig. \ref{fig:square}), this is an unusual situation in which the
\textbf{I} corner must imply and entail the \textbf{A} corner, 
in the particular case when $\phi$ is $\exq x G(x)$.  
G\"odel's proof shows that his axioms are indeed 
strong enough to invert the direction of
entailment for this choice of $\phi$.  
This observation, however, immediately leads 
to the question whether the axioms are eventually 
even strong enough to enable the inverted entailment
for any arbitrary sentence $\phi$.  
That is essentially the question asked by 
Sobel \cite{Sobel1987}, and his proof of the modal collapse
(\textbf{MC}, cf. Fig.~\ref{fig:collapse}) 
provides an affirmative answer. It is possible to show
that this form of the modal collapse entails (in modal logic
\textbf{\emph{K}}) a collapse of the modal square (\textbf{MCs}),
causing the subcontraries to entail 
(and even imply) their respective
contraries. Normally, as shown in Fig. \ref{fig:square}, 
in the modal square of opposition only the 
other direction of entailment holds: the
contraries entail their subcontraries, 
assuming the \emph{modal existential import} \textbf{ExImp}
\cite{ExImport}.

@@

Independently of the variant of the ontological argument 
under consideration, the following can be said about 
classes of collapsing propositions:
\begin{enumerate}
\item Valid propositions are collapsing: 
if $\phi$ is valid, then $\collapse{\phi}$ is valid.
%
\item The class of collapsing propositions 
is closed under logical equivalence: 
if $\collapse{\phi}$ is valid and $\phi \biimp \phi'$ is valid, 
then $\collapse{\phi'}$ is valid.
%
\item The class of collapsing propositions is 
not generally closed under equi-validity: 
even if $\collapse{\phi}$ is valid and $\phi$ and $\phi'$ 
are equi-valid, \\ $\collapse{\phi'}$ may not be valid.
%
\item The class of collapsing propositions is 
not generally closed under implication: 
even if $\collapse{\phi}$ is valid and 
$\phi \imp \phi'$ is valid, 
$\collapse{\phi'}$ may not be valid.
\end{enumerate}

### 2014-09-19 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Reorganize folders

Ontological arguments for the existence of God can be traced back at least to St. Anselm (1033-1109). His argument considered a maximally conceivable being, who must exist, because if it did not have the property of existence, then we could conceive of a greater being that, apart from the other properties, also has the property of existence.

A major critique of this argument is that we do not know whether the concept maximal conceivable being in fact designates anything or if it is inconsistent, like a round square \citep{fitting}[p.134].  
%The question is if the argument is sound. 
As Bertrand Russell pointed out, the definition of maximal allows us to define properties, like having boots, which the maximal being then also must have \citep{citation needed}.

@@

St. Anselm's argument was elaborated further by Descartes and Leibniz. Leibniz identified that establishing the possible existence of God is a critical missing step in St. Anselm's argument. To fill this gap, he argued that the properties of God, the perfections, are compatible. This implies that it is possible to have all perfections at once and therefore the existence of a maximal being with all these properties is possible.

@@

The first contribution of this paper is a detailed formalization of Scott's version \cite{Scott} of G\"odel's ontological argument \cite{Goedel} (as shown in Section \ref{sec:scott}) in a natural deduction calculus (as defined in Section \ref{sec:calculus}). A natural deduction style \citep{gentzen,prawitz} was chosen mainly for three reasons. Firstly, presentations of G\"oedel's proof are typically either informal or formalized in other styles of calculi (e.g. Fitting's tableaux \cite{fitting} or Sobel's sentential modal calculus \cite{sobel2}). Therefore, a formalization in natural deduction is a valuable complement to the existing presentations. Secondly, it makes the ontological proof accessible to people who are familiarized with a natural deduction style. Finally, as natural deduction is the style used by proof assistants such as Coq \cite{coq} and Isabelle \cite{isabelle}, the natural deduction formalizations have been easily checked step-by-step in Coq \cite{coqformalizations}.

@@

@@

>>>>

Ontological arguments for the existence of God can be traced back at least to St. Anselm (1033-1109). His argument considered a greatest conceivable being, who must exist, because if it did not have the property of existence, then we could conceive of a greater being that, apart from the other properties, also has the property of existence.

A major critique of this argument is that we do not know whether the concept greatest conceivable being in fact designates anything or if it is inconsistent, like a round square \citep{fitting}[p.134].  
%The question is if the argument is sound. 
As Bertrand Russell pointed out, the definition of greatest allows us to define properties, like having boots, which the greatest being then also must have \citep{citation needed}.

@@

St. Anselm's argument was elaborated further by Descartes and Leibniz. Leibniz identified that establishing the possible existence of God is a critical missing step in St. Anselm's argument. To fill this gap, he argued that the properties of God, the perfections, are compatible. This implies that it is possible to have all perfections at once and therefore the existence of a greatest being with all these properties is possible.

@@

The first contribution of this paper is a detailed formalization of Scott's version \cite{Scott} of G\"odel's ontological argument \cite{Goedel} (as shown in Section \ref{sec:scott}) in a natural deduction calculus (as defined in Section \ref{sec:calculus}). A natural deduction style \citep{gentzen,prawitz} was chosen mainly for three reasons. Firstly, presentations of G\"odel's proof are typically either informal or formalized in other styles of calculi (e.g. Fitting's tableaux \cite{fitting} or Sobel's sentential modal calculus \cite{sobel2}). Therefore, a formalization in natural deduction is a valuable complement to the existing presentations. Secondly, it makes the ontological proof accessible to people who are familiarized with a natural deduction style. Finally, as natural deduction is the style used by proof assistants such as Coq \cite{coq} and Isabelle \cite{isabelle}, the natural deduction formalizations have been easily checked step-by-step in Coq \cite{coqformalizations}.

@@

ToDo: decide which connectives and corresponding inference rules will be primitive.

@@

ToDo: decide if rules for diamond should be added.
